---
title: "Treatment Research"
description: |
  Exploring the value of research through two OCD treatment options
site: distill::distill_website
---

Obsessive-Compulsive Disorder (OCD) is one of the top ten most debilitating disorders, yet is 
rarely treated with the right medication.^[Source: The World Health Organization: The Global Burden of Disease]

OCD procedure front-line procedure is to prescribe Selective-Serotonin Reuptake Inhibitors (SSRIs),
and provide the patient with the option for Cognitive Behavioral Therapy (CBT). One of the most
common plans for CBT is to use Exposure-Response Prevention (ERP) over several months to train the 
patient to adapt to the disorder. This has proven to be semi-effective, but is quite expensive.^[Richards DA, Ekers D, McMillan D, Taylor RS, Byford S, Warren FC, Barrett B, Farrand PA, Gilbody S, Kuyken W, O'Mahen H, Watkins ER, Wright KA, Hollon SD, Reed N, Rhodes S, Fletcher E, Finning K. Cost and Outcome of Behavioural Activation versus Cognitive Behavioural Therapy for Depression (COBRA): a randomised, controlled, non-inferiority trial. Lancet. 2016 Aug 27;388(10047):871-80. doi: 10.1016/S0140-6736(16)31140-0. Epub 2016 Jul 23. PMID: 27461440; PMCID: PMC5007415.]

However, new studies have shown the effectiveness of the Bergen 4 Day Treatment, which instead of
months of costly CBT, compresses the ERP treatment into a 4-day intensive in-patient program that 
is not only more affordable, but also much more convenient and therefore less likely to be turned
down by patients.





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.height = 10, fig.width=15)
library(tidyverse)
library(readr)
library(stringr)
library(tidyr)
```

```{r}
table1 <- read.csv('/Users/isaacfrank/Documents/R Projects/OCDProject/T0004-10.1080_16506073.2018.1478447.csv')
table1 <- table1 %>%
  mutate(B4DT = NULL)  %>%
  rename(Measurement = X.) %>%
  mutate(ERP = NULL) %>%
  rename(B4DT = X) %>%
  rename(ERP = X.1) 
view(table1)

view(table1)

clean_data <- table1[-c(1,3,5,10), ]
clean_data$B4DT<-gsub("%","",as.character(clean_data$B4DT))
clean_data$ERP<-gsub("%","",as.character(clean_data$ERP))
clean_data$ERP<-gsub("\\*","",as.character(clean_data$ERP))
view(clean_data)

sorted <- levels(clean_data$Measurement) <- c("Declined Treatment", 
  "Dropped Treatment", "Remitted Post-assessment", "Not Remitted", "Relapsed",
  "Recovered (non-remitters)") %>% as_tibble() %>%
  mutate(B4DT = as.numeric(clean_data$B4DT), ERP = as.numeric(clean_data$ERP))
view(sorted)


final_data <- pivot_longer(sorted, cols = 2:3, names_to = "Treatment", values_to = "Proportion") %>%
  rename(Measurement = value) %>%
  group_by(Proportion)
view(final_data)
```

```{r}
final_data %>%
  ggplot(data = final_data, 
         mapping = aes(x = Measurement, y = Proportion, fill = Treatment, group = Treatment)) +
    geom_col( width = 0.8, position = position_dodge2(preserve = "single", reverse = TRUE)) + 
    coord_flip() +
    labs(title = "The Bergen 4 Day Treatment vs Ordinary ERP",
         x = "",
         y = "Proportion out of 100 patients",
         subtitle = "B4DT appears to be more effective, but possibly more temporary",
         caption = "Source: B4DT for OCD: 4 year follow-up of ERP, CBT (2019) Bjarne Hansen et al.") +
      theme(text = element_text(size = 20))
```



The Bergen 4 Day Treatment (B4DT) had a much higher remission rate, and much higher recovery 
rate than ordinary ERP. Although ordinary ERP had a lower chance of relapse, the difference is 
negligible when comparing to the more effective results from the B4DT. Additionally, because 
of how short the B4DT is, much fewer patients declined or dropped treatment when compared to
standard ERP procedure.

